\section{Introduction}
Software updates are everywhere. The most vital aspect for the sustainability of any software system is its ability to effectively and swiftly adapt to changes (i.e., to software updates). Therefore the adoption of changes is in the heart of the lifecycle of any system and blockchain systems are no exception. Software updates might be triggered by a plethora of different reasons. 

Typically, the main driver for a change will be a change request, or a new feature request by the user community. These type of changes will be planned for inclusion in some future release and then implemented and properly tested prior to deployment into production. In addition, another major source of changes is the correction of bugs (i.e., errors) and the application of security fixes. These usually produce the so-called \enquote{hot-fixes}, which are handled in a totally different manner (i.e., with a different process that results into different deployment speed) than change requests. In addition, depending on the severity level of the problem, or the priority (i.e., importance) of the change request there are different levels of speed and methods for the deployment of the software update into production. 

More specifically, for blockchain systems, a typical source of change are the enhancements at the consensus protocol level. There might be changes to the values of specific parameters (e.g., the maximum block size, or the maximum transaction size etc.), changes to the validation rules at any level (transaction, block, or blockchain), or even changes at the consensus protocol itself. Usually, the reason for such changes is the reinforcement of the protocol against a broader scope of adversary attacks, or the optimization of some aspect of the system like the transaction throughput, or the storage cost etc.

Finally, there are also more radical changes, which are usually caused by the introduction of new research ideas  and the advent of new technology, which becomes relevant. These type of changes usually introduce new concepts and are not just enhancements and thus trigger a major change to the system.

\paragraph{Context of this paper.} 
In this paper, our focus is on the software update mechanism of permission-less stake-based blockchain systems. We depart from the traditional centralized approach of handling software updates, which is the norm today for many systems (even for the ones that are natively decentralized, like permission-less blockchain systems) and try to tackle common software update challenges in a decentralized setting. We consider the full lifecycle of a software update, from conception to activation and propose ''decentralized alternatives'' to all phases. Essentially, we introduce a decentralized \emph{maintenance} approach for permission-less blockchain systems.

%\paragraph{Context of this paper.} In this paper, our focus is on the update mechanism of permissionless stake-based blockchain systems. We try to overcome known shortcomings of today's update methods for blockchain systems and propose a logical architecture for an update mechanism for stake-based ledgers. Our architecture covers all the components of a typical blockchain system and demonstrates how can such a mechanism be smoothly incorporated into each layer of the blockchain system. We describe in detail the various enhancements and changes in the existing components and then proceed into describing new components such as the governernance component, or the update-logic component. Our aim is to cover all aspects of software updating, from update proposal submission to update deployment and activation. Finally, we implement our logical architecture into a prototype update system for the Cardano blockchain. We then discuss various implementation issues and the prototype architecture.


\paragraph{Problem Definition.} 
Traditionally, software updates for blockchain systems have been handled in an ad-hoc, centralized manner: Somebody, often a trusted authority, or the original author of the software, provides a new version of the software, and users
download and install it from that authority's website. Even if the system follows an open source software development model \nnote{citation ?}, and therefore an update can potentially be implemented by anyone, the final decision of accepting or rejecting a new piece of code is always taken by the main maintainer (or group of core maintainers) of the system, who essentially constitutes a central authority. Even in the case, where the community has initially reached at a consensus for an update proposal (in the form of an \emph{Improvement Proposal} document \nnote{citation?}), through the discussion that takes place in various discussion-forums, still it remains an informal, ''social'' consensus, which is not recorded as an immutable historical event in the blockchain and the final decision is always up to the code maintainer to accept it, or not. Moreover, the authenticity and safety of the downloaded software is usually verified by the digital signature of a trusted authority, such as the original author of the software.

On the other hand, in a decentralized setting software updates are handled quite differently: An update proposal can be submitted by anyone (just like anyone can potentially submit a transaction to the blockchain). The decision of which update proposal will be applied and which wont, is taken collectively by the community and not centrally. Thus, the road-map of the system is decided jointly. However, this process is not an informal discussion process but it is part of an update protocol and all relevant events generated, are stored within the blockchain itself, and thus recorded in the immutable update history of the system. Moreover, in the decentralized setting, the role of the code maintainer, who takes decisions on the correctness of the submitted new code, the versions of the software and also guarantees the validity of the downloaded software, is replaced by the honest stake majority. 

Therefore, in the context of software updates for public stake-based blockchain systems, we define the capability to take decisions on: a) the priorities of software updates, b) the correctness of the new code, c) the maintenance of the code-base and d) the authenticity and safety of the produced binaries, based on honest stake majority decisions, which are part of the consensus protocol and recorded on-chain, while at the same time fulfilling software dependencies requirements, resolving version conflicts and enabling different update policies based on update metadata, as the \emph{decentralized software updates problem} and this is the problem that we are trying to solve in the paper.


%\paragraph{Problem Definition.}  Traditionally, software updates have been handled in an ad-hoc, centralized manner: Somebody, often
%a trusted authority, or the original author of the software, provides a new version of the software, and users
%download and install it from that authority's website. However, this approach is clearly not decentralised, and
%hence jeopardizes the decentralized nature of the whole system: In a decentralized software update mechanism,
%proposed updates can be submitted by anyone (just like anyone can potentially create a transaction in a blockchain).
%The decision of which update proposal will be applied and which wont, is taken collectively by the community
%and not centrally. Thus, the road-map of the system is decided jointly. Moreover, there is no central code
%repository, nor there is some main software maintainer, who decides on the code. All versions of the code
%are distributed and only local copies exist, in the same manner that the ledger of transactions is distributed in
%blockchain. The decentralized software update system must reach a consensus, as to what version is the current
%one (main branch) and which code branch will be merged into main. Finally, the decentralized software update
%system guarantees the authenticity and safety of the downloaded software, without the need to have some central authority
%to sign the code, in order to be trusted


%The traditional way of handling software updates is neither decentralized nor secure
%\begin{itemize}
%\item No standard way to propose updates
%\item Not a decentralized and democratic way to reach at a consensus on update priorities. Only \enquote{social consensus} is reached via social media. This is unstable and prone to chain splits.
%\item No essential auditing and verifiability of the agreement
%\item No standard way to record the immutable history of all these events (proposals,agreement, \item activation of updates)
%\item No standard method for security guarantees for the software installed
%\item No standard way to resolve conflicts and respect dependencies 
%\item No standard Update metadata
%\item One-size-fits-all for all \enquote{types} of changes (bugs,CRs)
%\item Hard-forks are the norm
%\end{itemize}

\import{./}{related_work.tex}

\paragraph{Goal of the paper.}
Our primary goal in this paper, is to propose a secure software update mechanism that will enable a decentralized approach to the blockchain software updates problem. We want to examine all phases in the lifecycle of a software update and propose decentralized alternatives that are practical and can be adopted in the real world. To this end, we investigate the smooth incorporation of such a decentralized software update mechanism into the architecture of a blockchain system and implement a prototype that realizes our ideas in practice. From a security perspective, our proposal must ensure that:
%...A safe and secure decentralized software update system. %In other words, a system that will ensure that: 
a) any honest stakeholder will always be able to submit an update proposal to be voted by the stakeholders community, b) it will prevent any update proposal that is not approved by the honest majority to be applied and c) it will provide guarantees for the authenticity and safety of the downloaded software.

\paragraph{Outline of the paper.}
%\noindent\textbf{Outline of the paper}