\begin{abstract}
%\subsection*{Abstract}
Software updates are a synonym to software evolution and thus are ubiquitous 
and inevitable to any blockchain platform. In this paper, we propose a general 
framework for decentralized software updates in distributed ledger systems. Our 
framework is primarily focused on Proof of Stake blockchains and aims at 
providing a solid set of enhancements, covering the full spectrum of a 
blockchain system, in order to ensure a decentralized, but also secure update 
mechanism for a public ledger. 
We identify what are the critical decisions in the lifecycle of a software 
update
and then propose a secure software update protocol that covers the \emph{full 	
lifecycle} of a software update from the ideation phase (the moment in which a 
change to the blockchain protocol is proposed) to the actual activation of the 
updated blockchain protocol, which enables decentralized decision making for 
\emph{all} critical decisions. We propose a liquid democracy scheme based on 
\emph{experts} for all the critical, but also deeply technical, decisions for 
software updates. We formally define what it means for a \emph{decentralized} 
software update system to be secure and propose secure activation protocols 
\cite{secure_activation} with various trade-offs. We deal with the complexity 
of priorities, version dependencies, conflicts resolution and emergency 
handling for the 
activation of updates and propose an elegant design. We perform voting and 
activation
threshold analysis, in order to achieve both properties of \emph{safety}
and \emph{liveness}, which we define in detail. Our performance analysis has 
shown 
that our update protocol is both linearly scalable in the number of 
participants and does not impact negatively the performance of the underlying 
blockchain. Finally, we implement our ideas into a research prototype and 
propose an architecture for integrating an update mechanism within the Cardano 
node \cite{cardano}; as well as discuss prototype validation with a 
trace-enabled 
property-based 
testing framework that we have developed. To the best of our knowledge, this is 
the first work that aims at formalizing the notion of a decentralized secure 
update for a blockchain and also that takes such a holistic approach on 
software updates.
\end{abstract}
