\section{Requirements for the software update mechanism} \label{sec:requirements}

\subsection{Functional requirements} \label{sec:func-reqs}
\begin{enumerate}
\item \textbf{Updates are open to the community.} \label{req:upd-open} Any
  stakeholder should be able to propose a software update to be voted by the
  stakeholders' community.

\item \textbf{Updates utilize the blockchain} \label{req:upd-blockchain} Update
  events are stored in the blockchain as immutable events.

\item \textbf{Decentralized Governance} \label{req-dec-gov} The stakeholders'
  community is the only authority to decide on the approval, or rejection, of
  the proposed software updates.

\item \textbf{Secure Activation} \label{req-sec-act} Upon activation of a
  software update the update mechanism guarantees a secure transition to the
  updated ledger; a transition in which:
  \begin{itemize}
  \item The security assumptions of the updated ledger will hold (e.g., there
    will be a sufficient percent of upgraded honest stake)
  \item The updated ledger successfully incorporates the old ledger, i.e., the
    state of the old ledger will be successfully moved into the new ledger.
  \item There will be no chain splits due to the activation.
\end{itemize}

\item \textbf{Conflict Resolution.}\label{req:conflict-res} Two or more software
  updates that have conflicting changes will not be applied both, but a conflict
  resolution will take place.

\item \textbf{Update Dependencies.}\label{req:update-dep} A software update will
  never be applied if its dependency requirements are not met

\item \textbf{Extensible voting period.} \label{req:ext-vper} The period
  required for approving a software update is not fixed but can be extended by
  the stakeholders' community. In other words, revoting is supported.

\end{enumerate}

\subsection{Non-functional requirements} \label{sec:non-func-reqs}

\begin{enumerate}
\item \textbf{Transaction Throughput}
  \label{req:tr-throughput}
  The ledger transaction throughput is not significantly impacted by the
  presence of software updates (see section \ref{sec:measurements}).

\item \textbf{Blockchain Size} \label{req:blockchain-size} The blockchain size
  is not significantly impacted by the presence of software updates (see section
  \ref{sec:measurements}).

\item \textbf{Update Time to Activation}\label{req:update-thr} The time to
  activation of a software update reduces linearly in the number of users
  running the update protocol (see section \ref{sec:measurements}).

\item \textbf{Processing Time}
  \label{req:proc-time}
  The processing time of a node due to the execution of the update protocol is
  not significantly increased.

\end{enumerate}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "decentralized-updates"
%%% End:
