\paragraph{Related Work.}
%\mnote{We should stress that this is the first paper that tries to formalize the problem of updates for blockchain. I think that this is the most novel thing that we have.}
To the best of our knowledge, there is no related work on the problem of the
decentralization of software updates in the context of blockchain systems in a
holistic manner, i.e., taking into consideration all phases in the lifecycle of
a software update.
Bitcoin \cite{bitcoin}, Bitcoin Cash \cite{bitcoincash}, Ethereum
\cite{ethereum} and Zcash \cite{zerocash} use a \say{social governance} scheme,
in which decisions on update proposals is reached through discussions on social
media. This type of informal guidance is  too unstable and prone to chain
splits, or prone to becoming too de-facto centralized \cite{buterin_on_gov}.
There exist blockchain systems \cite{dash}, \cite{decred} that adopt a
decentralized governance scheme, in which the priorities, as well as the
funding of update proposals is voted on-chain as part of a maintenance
protocol. However, these proposals do not follow a holistic approach to the
decentralization software updates problem. Instead, their focus is merely on
the ideation phase in the lifecycle of a software update, where an update
proposal is born as an idea, and the community is called to accept if it will
be funded or rejected. These solutions do not deal with the remaining phases in
this lifecycle, which pertain to the approval of the source code correctness
%and the authenticity of the binaries that will be distributed for downloading,
%the maintenance of the code base
and more importantly, with the activation of the changes. That is why in the
above cases, there exists a central authority (or group) that assumes the role
of the source code maintainer. A similar approach is proposed by Bingsheng et
al. \cite{treasury}, where a complete treasury system is proposed for
blockchain systems, in which liquid democracy / delegative voting is followed.
We also follow the approach of voting delegation, when technical expertise is
required in order to reach a decision for an update proposal. Similarly,
Bingsheng et al. work is focused only on the treasury system, which covers only
the initial phase in the lifecycle of an update proposal.


%Good examples of governance:
%\begin{itemize}
%\item Treasury paper (https://eprint.iacr.org/2018/435)
%\item Dash Governance  - https://www.dash.org/governance/
%\item Decred Governance - https://docs.decred.org/governance/consensus-rules-voting/
%\end{itemize}
%
%Bitcoin and Ethereum use a \enquote{Social Governance} scheme, which basically means that decisions on update proposals is reached through discussions on social media. This type of informal guidance is  too unstable and prone to chain splits, or prone to becoming too de-facto centralized [https://vitalik.ca/general/2017/12/17/voting.html]
